\Huge Conferences in an Era of Expensive Carbon

### 2019-06-19 main.tex
Checkpoint

In fact, scientific societies arguably have an especially important role to
play, since, for many of their members, travel to conferences may be a
substantial or even dominant part of their individual contribution to
climate change.  A single round-trip flight from Philadelphia to Paris emits
the equivalent of about 1.7 tons of carbon dioxide (\COtwoE) per
passenger~\cite{carboncalculator}\iflater\bcp{Trim citation?}\fi---a
significant fraction of the {\em total} yearly emissions for an average
resident of the US (16.5 tons) or Europe (7 tons)~\cite{emissions}.
\iflater\bcp{For comparison, ...  (put in some figures about the footprints
  of a few other things that people might think of reducing---driving,
  eating meat, upgrading their fridge...)}\fi{} Moreover, these emissions
have no near-term technological fix, since jet fuel is difficult to replace
with renewable energy sources~\cite{elec-air}.
% \iflater\bcp{Trimmed this sentence: Reducing the \COtwoE{} footprint
% of conferences will require a reduction in the footprint of the attendees'
% air travel.}\fi

How should ACM respond to these facts?  We have two concrete
proposals. First, all ACM conferences should publicly account for CO2e
spent as a result of hosting them. Second, ACM should embrace putting
a \emph{price} on carbon in conference budgets, to create pressure on
organizers to reduce its consumption.

\SECTION{Mandate public accounting of carbon footprints.}  
\bcp{The section headers could use some wordsmithing!}

Our first proposal is simple: Every ACM-sponsored conference should
publicly report its carbon footprint, total and per-participant. A
conference's footprint will be dominated by participants' air travel,
but the accounting should not be limited to that: It should also
include air travel to in-person PC meetings, and estimate hotel and
food CO2e costs.

We hope that a public accounting will nudge conference organizers and
attendees to change their behavior. As an analogy: Chain restaurants
in the US are required by law to post calorie counts of food items on
menu boards, and studies show that enlightened customers order, on
average, up to 50 fewer calories a day, as a
result~\cite{menu}.

ACM can provide tools to help organizers account for this
footprint. We recently worked with the ACM SGB to develop an air
travel-focused CO2e calculator. One can input a CSV with registration
data, and the calculator will compute the estimated CO2e cost of air
travel for those sufficiently far away. This calculator could be
expanded to support related costs.

SIGPLAN has been thinking about how to reduce CO2e usage over the past
couple of years, informed by an accounting of CO2e expenditures. Doing
so has led several conference organizers \mwh{name which conferences}
to switch from holding in-person PC meetings to holding the meetings
on-line instead. \mwh{Mention Jens survey here.} It has also prompted
adding support for conference livestreaming (to go with talk
recording) to permit remote participation.

@@

This dilemma is the one that all of society faces, in a microcosm. To
address it, many policy experts advocate some form of carbon tax (or
\emph{price}), which adds a near-term, monetary cost to greenhouse gas
emissions~\cite{carbonprice}. Doing so makes manifest today the hidden
environmental cost of CO2e emissions on tomorrow, thus incentivizing
CO2e-reducing changes without specifying exactly which, thus allowing
for creative and efficient responses. As an analogy, some
municipalities such as New York city and Philadelphia have imposed a
per-calorie tax on soft drinks, with some studies finding that doing
so indeed reduced consumption~\cite{sodatax}.

In the best case, governments would impose a CO2e tax uniformly, and
all CO2e-intensive activities would have to pay it. But ACM could send
a strong message about the importance of this issue by acting now. In
particular, as several companies have shown, even an artificial ``internal price'' on
  carbon creates a significant price signal that shapes behavior.
%
  Thus, our second (more ambitious) proposal is: ACM should begin requiring
  conferences to pay a surcharge based on their carbon footprint.  It should
  be low at the beginning and increase steadily (and predictably) year
  on year.
%
Conference organizers can choose (on the basis of
  careful analysis of data and weighing of benefits) how best to do
  so.

 What should ACM do with the funds collected? The money raised in
  this way can be used in many ways: buying carbon offsets, supporting
  ``green computing research'', virtualizing conferences
  (livestreaming, etc.), supporting a transition to a low-carbon / few
  conferences business model for ACM.

Once again, ACM should provide support to conference organizers to
prepare for a future of higher-cost conferences.

@@

(Some kind of closing.)

>>>>

Indeed, scientific societies arguably have an especially important role to
play, since, for many of their members, travel to conferences may be a
substantial or even dominant part of their individual contribution to
climate change.  A single round-trip flight from Philadelphia to Paris emits
the equivalent of about 1.7 tons of carbon dioxide (\COtwoE) per
passenger~\cite{carboncalculator}\iflater\bcp{Trim citation?}\fi---a
significant fraction of the {\em total} yearly emissions for an average
resident of the US (16.5 tons) or Europe (7 tons)~\cite{emissions}.
\iflater\bcp{For comparison, ...  (put in some figures about the footprints
  of a few other things that people might think of reducing---driving,
  eating meat, upgrading their fridge...)}\fi{} Moreover, these emissions
have no near-term technological fix, since jet fuel is difficult to replace
with renewable energy sources~\cite{elec-air}.
% \iflater\bcp{Trimmed this sentence: Reducing the \COtwoE{} footprint
% of conferences will require a reduction in the footprint of the attendees'
% air travel.}\fi

How should ACM respond to these facts?  We have two concrete
proposals. First, all ACM conferences should publicly {\em account} for CO2e
emitted as a result of putting them on. Second, ACM should put a
\emph{price} on carbon in conference budgets, to create pressure on
organizers to reduce their footprints.

\SECTION{Mandate public accountability.}  
% \bcp{The section headers could use some wordsmithing!}

Our first proposal is simple: Every ACM-sponsored conference should
calculate and publicly report its carbon footprint (both aggregate and
per-participant). Most conferences' footprints will be dominated by
participants air travel, but the accounting should go beyond this to include
air travel to in-person PC meetings and estimated hotel and food CO2e
footprints.

ACM can provide tools to measure emissions. As a start, SIGPLAN (the special
interest group on programming languages) recently worked with ACM's SIG
Governing Board to develop an air-travel-focused CO2e calculator for
conferences. Users can upload conference registration data, and the
calculator will estimate the CO2e cost of air travel for those sufficiently
far away. This calculator could be expanded to account for other conference
emissions.

Public accounting will nudge conference organizers and attendees to change
their behavior. By analogy, chain restaurants in the US are required by law
to post calorie counts of food items on menu boards, and studies show that
enlightened customers order, on average, up to 50 fewer calories a day as a
result~\cite{menu}.

SIGPLAN has been considering how to reduce CO2e usage over the past couple
of years, informed by an accounting of its CO2e expenditures. This
discussion has led the organizers of some of its flagship
conferences\mwh{name which conferences}\bcp{POPL and ICFP, not PLDI; I don't
  remember where SPLASH stands} to switch from in-person to on-line program
committee meetings, joining a trend among other SIGs. \mwh{Mention Jens
  survey here.} It has also prompted conferences to increase their
investments in livestreaming and talk recording to permit remote
participation.

@@

This dilemma is a microcosm of the one that all of society faces. To address
it, many policy experts advocate some form of ``carbon pricing'' (either
taxes or cap-and-trade schemes), which imposes a concrete, immediate cost on
greenhouse gas emissions~\cite{carbonprice}. Doing so makes manifest the
hidden environmental cost of CO2e emissions, thus incentivizing
CO2e-reducing changes without specifying exactly which, allowing for
creative and efficient responses. (Continuing the junk-food analogy, some
municipalities such as New York city and Philadelphia have imposed a
per-calorie tax on soft drinks, with some studies finding that doing so
indeed reduced consumption~\cite{sodatax}.)

Ideally, at some point, governments will impose carbon pricing uniformly,
and all CO2e-intensive activities will have to pay it. But ACM can send a
strong message about the importance of this issue by acting now. In
particular, as several companies have shown, even an artificial ``internal
price'' on carbon creates a significant price signal that shapes behavior.
%
Thus, our second (more ambitious) proposal is: ACM should impose a a
surcharge on conferences based on their carbon footprint.  The charge should
be low at the beginning and increase steadily (and predictably) year on
year.
%
Conference organizers can then choose (on the basis of careful analysis of
data and weighing of benefits) how best to balance their budgets---whether
by decreasing emissions, increasing registration fees, soliciting additional
corporate sponsorship, or some combination.

What should ACM do with the funds collected from this surcharge?  One can
imagine many good uses, including buying carbon
offsets~\cite{CarbonOFfsetReport}, supporting ``green'' computing
research\bcp{can we mention some specifics?}, and defraying the costs of
virtualizing conferences (livestreaming, etc.).

@@

The climate crisis is too important to ignore and too urgent to leave to
world leaders, to address at their own pace: rather, organizations at every
scale must confront their own contributions, raise awareness among their
members, and begin establishing new ways of doing business in the
lower-carbon future that is coming.  ACM should lead the way!
